Data Prep

### 2024-01-17 00.0_dataPrep.qmd
More explanation, remove lfr individuals later on in the pipeline

## Remove individuals with low fix rate

Some of the INPA vultures were tracked only once an hour, or some other low fix rate. However, determining which birds these are is tricky because even the vultures that should have a 10-minute fix rate do not always have points every 10 minutes, e.g. due to lack of reception or low tag battery.\
\
Therefore, we have to identify **low fix rate (lfr) individuals**, which are those that *never* have a 10-minute fix rate. To do this, we will find the mode (most common) fix frequency for each individual on each day. Then, we will exclude individuals who never have observations at 10-minute intervals. We will keep any that theoretically have a 10-minute fix rate (i.e. their observation frequency was 10 minutes at some point), even if in practice they often have much less frequent observations.

@@

-   `data/dataPrep/removed_lfr.Rda`

@@

# Exclude individuals that never have a daily mode of 10 minutes
toRemove_lfr <- map(with_modes, ~.x %>%
                        sf::st_drop_geometry() %>%
                        group_by(Nili_id) %>%
                        summarize(minmode = min(mode, na.rm = T)) %>%
                        filter(minmode > 10) %>%
                        pull(Nili_id) %>%
                        unique())

removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% 
                                   filter(!(Nili_id %in% .y)))

save(removed_lfr, 
     file = "data/dataPrep/removed_lfr.Rda")

# Tidy up
rm(seasons_list)
rm(removed_lfr)
rm(with_modes)
rm(toRemove_lfr)
gc()
```

@@

The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So when we do the social analysis, we will use `removed_lfr` and not anything further along.

@@

We are going to need to get nightly roost locations for all of the remaining vultures. We have to do this now because getting roost locations requires using points that occur at night, which we will eventually want to remove in order to do the movement analysis.

@@

-   `data/dataPrep/removed_lfr`

@@

```{r}
load("data/dataPrep/removed_lfr.Rda")
```

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = "data/dataPrep/roosts.Rda")

# Tidy up
rm(removed_lfr)
rm(roosts)
gc()
```

# Further processing for movement data

@@

## Remove individuals with too few points per day

@@

## Remove individuals with too few days tracked

>>>>

## Identify individuals with low fix rate

Some of the INPA vultures were tracked only once an hour, or some other low fix rate. However, determining which birds these are is tricky because even the vultures that should have a 10-minute fix rate do not always have points every 10 minutes, e.g. due to lack of reception or low tag battery.\
\
Therefore, we have to identify **low fix rate (lfr) individuals**, which are those that *never* have a 10-minute fix rate. To do this, we will find the mode (most common) fix frequency for each individual on each day. Then, we will identify individuals who never have observations at 10-minute intervals. We will keep any that theoretically have a 10-minute fix rate (i.e. their observation frequency was 10 minutes at some point), even if in practice they often have much less frequent observations.

Because tags can be put on and off and the fix rate can change, I'm calculating the fix rate on a per-season basis (which is why I didn't put this step earlier in the pipeline–it had to come after the season splitting).

I suppose we could leave in the low fix rate individuals, especially for co-roosting. \[Actually, maybe we should leave them in for co-roosting, but not the others... which makes me think maybe we should leave them in for roost and social, just not movement, because they can contribute to the background information. Aaaarrrrrrggghhhhh.... XXX\]

@@

-   `data/dataPrep/low_fix_rate_indivs.Rda`

```{r}
load("data/dataPrep/seasons_list.Rda")
load("data/dataPrep/season_names.Rda")
```

@@

# Identify individuals that never have a daily mode of 10 minutes
low_fix_rate_indivs <- map(with_modes, ~.x %>%
                        sf::st_drop_geometry() %>%
                        group_by(Nili_id) %>%
                        summarize(minmode = min(mode, na.rm = T)) %>%
                        filter(minmode > 10) %>%
                        pull(Nili_id) %>%
                        unique())
names(low_fix_rate_indivs) <- season_names

save(low_fix_rate_indivs, 
     file = "data/dataPrep/low_fix_rate_indivs.Rda")
# Don't need to save with_modes because we've already used this step to identify low fix rate individuals.

# Tidy up
rm(seasons_list)
rm(season_names)
rm(low_fix_rate_indivs)
rm(with_modes)
gc()
```

@@

The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So when we do the social analysis, we will use `seasons_list` and not anything further along.

@@

We are going to need to get nightly roost locations for all of the vultures. We have to do this now because getting roost locations requires using points that occur at night, which we will eventually want to remove in order to do the movement analysis.

@@

@@

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(seasons_list, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = "data/dataPrep/roosts.Rda")

# Tidy up
rm(seasons_list)
rm(roosts)
gc()
```

# Further processing for movement analysis

## Remove low fix rate individuals

-   `data/dataPrep/low_fix_rate_indivs.Rda`

```{r}
load("data/dataPrep/seasons_list.Rda")
load("data/dataPrep/low_fix_rate_indivs.Rda")
```

```{r}
removed_lfr <- map2(seasons_list, low_fix_rate_indivs,
                    ~.x %>% filter(!(Nili_id %in% .y)))
save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")

# tidy up
rm(seasons_list)
rm(low_fix_rate_indivs)
rm(removed_lfr)
gc()
```

@@

## Remove vulture-days with too few points per day

@@

## Remove individuals with too few high-ppd days tracked
